R tools for accessing research literature for text mining

### 2019-04-16 manuscript.Rmd
updates

There's more than 100 million research articles published
(Crossref API: <https://github.com/CrossRef/rest-api-doc>), representing
an enormous amount of knowledge.  In addition to simply reading these
articles, the articles contain a vast trove of information of interest to
researchers for machine aided questions [@Kong2018; @Usai2018].

For example, many researchers are interested in statistical
outcomes of articles: questions about P-values, about effect
sizes, and more. With regard to effect sizes, these are
of particular interest, as they are often combined in
meta-analyses to draw broad conclusions about a particular
question.

Text-mining is the broad term associated with pulling information
out of articles. Given the importance of text-mining, good
text-mining tools are needed to make it easier for researchers to do.
Graphical user interface (GUI) based text mining tools are available
[e.g., @ba2016interoperability] and some research papers have used them
[@Chaix2019], but given the urgent recent call to action for more reproducible
research [@OSC2015; @Camerer2016; @Camerer2018],
we must move away from GUI based tools as fast as possible.

In particular, the R programming language is used widely throughout many
academic fields and thus tools in R for text mining are of particular
importance because they can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those
tools for searching for, acquiring, and extracting particular chunks
of texts (e.g., title, abstract, authors). Most of the packages are
part of the rOpenSci suite (https://ropensci.org/).

@@

A text-mining project starts with a question. From that question,
researchers then attempt to acquire scholarly articles for
text-mining. Finding those articles is not altogether straight-forward.

There are many places to search for articles; a non-exhaustive list:
Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect,
Web of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
Open Knowledge Maps, and more. It's probably difficult to know where the
best place is to search. Some of these are paywalled (e.g., Web of Science),
and some are not.

@@

@@

Data Provider                      | Content Type            | rOpenSci Package         | Documentation
---------------------------------- | ---------------------   | ------------------------ | -----------------
Crossref                           | Metadata only           | rcrossref/crminer        | [^1]
DataCite                           | Metadata only           | rdatacite                | [^2]
Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^11]
Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
JSTOR Data for Research            | Full content            | jstor                    | [^13]
ORCID                              | Metadata                | rorcid                   | [^14]
Wikimedia's Citoid                 | Citations               | rcitoid                  | [^15]
Open Citation Corpus               | Citations               | citecorp                 | [^16]

[^1]: https://api.crossref.org
[^2]: https://support.datacite.org/docs/api
[^3]: http://bit.ly/KYQ1Rd
[^4]: http://api.plos.org/solr
[^5]: http://bit.ly/J9S616
[^6]: https://arxiv.org/help/api/index
[^7]: https://dev.springer.com/
[^8]: http://www.biorxiv.org/
[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
[^10]: https://azure.microsoft.com/en-us/services/cognitive-services
[^11]: https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97
[^12]: https://doaj.org/api/v1/docs
[^13]: https://www.jstor.org/dfr/
[^14]: https://pub.orcid.org/
[^15]: https://en.wikipedia.org/api/rest_v1/#/Citation/getCitation
[^16]: http://opencitations.net/

@@

# fulltext: a swiss army knife for text mining in R

[fulltext][] is a general purpose R package for the data part of text-mining:
search for articles, get links to articles, get article abstracts, and fetch
full text of articles. The `fulltext` package is always adding additional
data sources as time allows (See Table 1). Starting from searching for
articles, the outputs of search can be fed into a function to get links
to those articles, or to get abstracts for those articles, or to fetch
their full text.

The following is a breakdown of the major distinct parts of `fulltext`.

@@

To make text mining easier:

1. publishers should provide XML if they have it
2. publishers should not change URL patterns so often, or at all
3. publishers should keep their Crossref metadata up to date
4. remove all paywalls (Easy, yes?)
5. ...

@@

[rcrossref]: https://github.com/ropensci/rcrossref
[rorcid]: https://github.com/ropensci/rorcid
[fulltext]: https://github.com/ropensci/fulltext
[microdemic]: https://github.com/ropensci/microdemic
[arXiv]: https://github.com/ropensci/arXiv
[rdatacite]: https://github.com/ropensci/rdatacite
[rbhl]: https://github.com/ropensci/rbhl
[rentrez]: https://github.com/ropensci/rentrez
[rplos]: https://github.com/ropensci/rplos
[crminer]: https://github.com/ropensci/crminer
[europepmc]: https://github.com/ropensci/europepmc
[jaod]: https://github.com/ropensci/jaod
[jstor]: https://github.com/ropensci/jstor
[rcitoid]: https://github.com/ropenscilabs/rcitoid
[citecorp]: https://github.com/ropenscilabs/citecorp
[serrano]: https://github.com/sckott/serrano
[habanero]: https://github.com/sckott/habanero
[quanteda]: https://quanteda.io/
[pubchunks]: https://github.com/ropensci/pubchunks
[pdftools]: https://github.com/ropensci/pdftools
[libpoppler]: https://poppler.freedesktop.org/
[tesseract]: https://github.com/ropensci/tesseract

>>>>

There's more than 100 million research articles published
(Crossref API: <https://github.com/CrossRef/rest-api-doc>), representing
an enormous amount of knowledge. In addition to simply reading these
articles, the articles contain a vast trove of information of interest to
researchers for machine aided questions [@Kong2018; @Usai2018]. For example,
many researchers are interested in statistical outcomes of articles that 
can be extracted from numeric results: P-values, effect sizes, means, and
more. In addition, researchers are often interested in words in articles,
their use through time, and the contexts they are found in.

Text-mining is the broad term associated with pulling information
out of articles. Given the importance of text-mining, good
text-mining tools are needed to make it easier for researchers to do.
Graphical user interface (GUI) based text mining tools are available
[e.g., @ba2016interoperability; @Munoz2019; @Canada2017] and some research papers
have used them [@Chaix2019], but given the urgent recent call to action for
more reproducible research [@OSC2015; @Camerer2016; @Camerer2018],
we must move away from GUI based tools as fast as possible. A number of examples
of programmatic tools can be found in the literature. For example, @Sinclair2016
present a tool in Python called seqenv for the domain specific task of
linking sequences to environments through text mining.

Most recent text-mining papers do not use programmatic approaches, highlighting
the need for more programmatic text-mining tools, and increased
discussion of those tools to increase awareness. For example, many papers 
search Web of Science using their web interface, and downloading papers manually 
[@Ding2018; @McCallen2019]. Many of these papers doing GUI based searching and 
paper downloads are using R or Python downstream for analysis; replacing 
GUI based data acquisition with programmatic approaches will improve 
research.

The R programming language is used widely throughout many academic fields;
tools in R for text mining are of particular importance because they
can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those
tools for searching for, acquiring, and extracting parts of texts
(e.g., title, abstract, authors). Most of the packages presented here are
part of the rOpenSci suite (https://ropensci.org/).

@@

A text-mining project starts with a question. From that question, researchers
then attempt to acquire scholarly articles for text-mining. Finding appropriate
articles is not altogether straight-forward.

Some of the discovery difficulty relates to the fact that there are so many
places to search for articles; a non-exhaustive list:
Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect,
Web of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
Open Knowledge Maps, CORE, Fatcat, and more. It's probably difficult to know
where the best place is to search. Some of these are paywalled
(e.g., Web of Science), and some are not.

@@

On top of the vast array of different data sources is the varied ways that 
search is implemented in each source. Most sources are probably using Solr
or Elasticsearch under the hood, though we can't know this for sure as most 
do not make their software infrastructure public knowledge. Nonetheless, data
sources differ in how search works from the user perspective. For example, some 
provide wild card search and some do not. Some sources are searching full text
of articles, while others only search metadata (i.e., title, authors, abstract).
In addition, each source has a different set of metadata/full text available.
In brief, the same search against different sources produces different results.
Some text-mining research articles perform the same search against many
different sources (refs), while others choose just one source.

@@

Data Provider                      | Content Type            | rOpenSci Package         | Documentation
---------------------------------- | ---------------------   | ------------------------ | -------------
Crossref                           | Metadata                | rcrossref/crminer        | [^1]
DataCite                           | Metadata                | rdatacite                | [^2]
Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
Public Library of Science (PLoS)   | Full content/altmetrics | rplos                    | [^4]
Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
Microsoft Academic Search          | Metadata                | microdemic               | [^11]
Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
JSTOR Data for Research            | Full content            | jstor                    | [^13]
ORCID                              | Metadata                | rorcid                   | [^14]
Wikimedia's Citoid                 | Citations               | rcitoid                  | [^15]
Open Citation Corpus               | Citations               | citecorp                 | [^16]
Fatcat                             | Metadata                | none                     | [^17]
SHERPA/RoMEO                       | Journal Level Metadata  | rromeo                   | [^18]
CORE                               | Full content/Metadata   | rcoreoa                  | [^19]
Dissemin                           | Metadata                | dissemr                  | [^20]

[^1]: https://api.crossref.org
[^2]: https://support.datacite.org/docs/api
[^3]: http://bit.ly/KYQ1Rd
[^4]: http://api.plos.org/solr
[^5]: http://bit.ly/J9S616
[^6]: https://arxiv.org/help/api/index
[^7]: https://dev.springer.com/
[^8]: http://www.biorxiv.org/
[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
[^10]: https://azure.microsoft.com/en-us/services/cognitive-services
[^11]: https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97
[^12]: https://doaj.org/api/v1/docs
[^13]: https://www.jstor.org/dfr/
[^14]: https://pub.orcid.org/
[^15]: https://en.wikipedia.org/api/rest_v1/#/Citation/getCitation
[^16]: http://opencitations.net/
[^17]: https://fatcat.wiki/
[^18]: http://www.sherpa.ac.uk/romeo/apimanual.php?la=en&fIDnum=|&mode=simple
[^19]: https://core.ac.uk/
[^20]: https://dissemin.readthedocs.io/en/latest/api.html

@@

## Fatcat

Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is a 
"versioned, publicly-editable catalog of research publications:  journal articles,
conference proceedings, pre-prints, blog posts". Fatcat is currently does not have an
R client, but is used inside of the [fulltext][] package.

## SHERPA/RoMEO

SHERPA/RoMEO (http://sherpa.mimas.ac.uk/romeo/index.php) aggregates and analyses
publisher open access policies and provides summaries of self-archiving permissions
and conditions of rights given to authors. The [rromeo][] is an R interface
to SHERPA/RoMEO.

## CORE

CORE (https://core.ac.uk/) touts itself as the world's largest collection of open
access research articles, providing metadata on journals and articles, as well
as access to the full text of articles. The [rcoreoa][] R package interfaces
with the CORE API.

## Dissemin

Dissemin (https://dissem.in/) detects papers behind pay-walls and invites authors
to upload them to an open repository. Dissemin provides metadata including links to
open versions of articles. The [dissemr][] R package interfaces with the
Dissemin API.

# fulltext: a toolset for text mining in R

[fulltext][] is a general purpose R package for the data part of text-mining:
search for articles, get links to articles, get article abstracts, and fetch
full text of articles. The `fulltext` package is always adding additional
data sources as time allows (See Table 1). Starting from searching for
articles, the outputs of search can be fed into a function to get links
to those articles, or to get abstracts for those articles, or to fetch
their full text. The following is a breakdown of the major distinct
parts of `fulltext`.

@@

Text-mining will always be a complex task given all the layers involved:
often temporal time-span of research questions; varied permissions among 
researchers and their articles they're trying to access; varied approaches
to getting full text (xml vs pdf vs plain text); and more.

Programmatic text-mining is a first step towards making text-mining easier.
The R ecosystem is an especially good place to do text-mining because 
there are many packages for text-mining analysis, and endless packages for
any required statistical analyses. In addition, rOpenSci and others are 
building up a set of packages in R for searching for and acquiring full
text programatically to help make the research workflow as reproducible
as possible.

Future work for `fulltext` includes:

1. Adding more publisher plugins
2. Fine tuned user control over publishers 
3. Improve VPN/proxy controls
4. Incorporate more search engines to help resolve URLs for fulltext versions
5. Improve documentation

With respect to what publishers can do to make text-mining easier, publishers
should:

1. provide XML if they have it
2. not change URL patterns so often, or at all
3. maintain consistent URL patterns among journals, years, etc.
4. keep their Crossref metadata up to date
5. open up their citation data

@@

[rcrossref]: https://github.com/ropensci/rcrossref
[rorcid]: https://github.com/ropensci/rorcid
[fulltext]: https://github.com/ropensci/fulltext
[microdemic]: https://github.com/ropensci/microdemic
[arXiv]: https://github.com/ropensci/arXiv
[rdatacite]: https://github.com/ropensci/rdatacite
[rbhl]: https://github.com/ropensci/rbhl
[rentrez]: https://github.com/ropensci/rentrez
[rplos]: https://github.com/ropensci/rplos
[crminer]: https://github.com/ropensci/crminer
[europepmc]: https://github.com/ropensci/europepmc
[jaod]: https://github.com/ropensci/jaod
[jstor]: https://github.com/ropensci/jstor
[rcitoid]: https://github.com/ropenscilabs/rcitoid
[citecorp]: https://github.com/ropenscilabs/citecorp
[serrano]: https://github.com/sckott/serrano
[habanero]: https://github.com/sckott/habanero
[quanteda]: https://quanteda.io/
[pubchunks]: https://github.com/ropensci/pubchunks
[pdftools]: https://github.com/ropensci/pdftools
[libpoppler]: https://poppler.freedesktop.org/
[tesseract]: https://github.com/ropensci/tesseract
[rcoreoa]: https://github.com/ropensci/rcoreoa

### 2019-04-16 manuscript.tex
updates

There's more than 100 million research articles published (Crossref API:
\url{https://github.com/CrossRef/rest-api-doc}), representing an
enormous amount of knowledge. In addition to simply reading these
articles, the articles contain a vast trove of information of interest
to researchers for machine aided questions (Kong \& Gerstein, 2018; Usai
et al., 2018).

For example, many researchers are interested in statistical outcomes of
articles: questions about P-values, about effect sizes, and more. With
regard to effect sizes, these are of particular interest, as they are
often combined in meta-analyses to draw broad conclusions about a
particular question.

Text-mining is the broad term associated with pulling information out of
articles. Given the importance of text-mining, good text-mining tools
are needed to make it easier for researchers to do. Graphical user
interface (GUI) based text mining tools are available (e.g., Ba \&
Bossy, 2016) and some research papers have used them (Chaix et al.,
2019), but given the urgent recent call to action for more reproducible
research (Open Science Collaboration, 2015; Camerer et al., 2016, 2018),
we must move away from GUI based tools as fast as possible.

In particular, the R programming language is used widely throughout many
academic fields and thus tools in R for text mining are of particular
importance because they can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those tools for
searching for, acquiring, and extracting particular chunks of texts
(e.g., title, abstract, authors). Most of the packages are part of the
rOpenSci suite (\url{https://ropensci.org/}).

@@

\begin{longtable}[]{@{}llll@{}}
\toprule
Data Provider & Content Type & rOpenSci Package &
Documentation\tabularnewline
\midrule
\endhead
Crossref & Metadata only & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
DataCite & Metadata only & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
Biodiversity Heritage Library & Full content/Metadata & rbhl &
\footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
Public Library of Science (PLoS) & Full text/altmetrics & rplos &
\footnote{\url{http://api.plos.org/solr}}\tabularnewline
Scopus (Elsevier) & Full content/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
arXiv & Full content/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
Biomed Central (via Springer) & Full content/Metadata & fulltext &
\footnote{\url{https://dev.springer.com/}}\tabularnewline
bioRxiv & Full content/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
PMC/Pubmed (via Entrez) & Full content/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
Europe PMC & Full content/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
Microsoft Academic Search & Metadata & fulltext/microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
JSTOR Data for Research & Full content & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
Wikimedia's Citoid & Citations & rcitoid & \footnote{\url{https://en.wikipedia.org/api/rest_v1/\#/Citation/getCitation}}\tabularnewline
Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{fulltext-a-swiss-army-knife-for-text-mining-in-r}{%
\section{fulltext: a swiss army knife for text mining in
R}\label{fulltext-a-swiss-army-knife-for-text-mining-in-r}}

\href{https://github.com/ropensci/fulltext}{fulltext} is a general
purpose R package for the data part of text-mining: search for articles,
get links to articles, get article abstracts, and fetch full text of
articles. The \texttt{fulltext} package is always adding additional data
sources as time allows (See Table 1). Starting from searching for
articles, the outputs of search can be fed into a function to get links
to those articles, or to get abstracts for those articles, or to fetch
their full text.

The following is a breakdown of the major distinct parts of
\texttt{fulltext}.

@@

To make text mining easier:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  publishers should provide XML if they have it
\item
  publishers should not change URL patterns so often, or at all
\item
  publishers should keep their Crossref metadata up to date
\item
  remove all paywalls (Easy, yes?)
\item
  \ldots{}
\end{enumerate}

@@

@@

@@

@@

>>>>

There's more than 100 million research articles published (Crossref API:
\url{https://github.com/CrossRef/rest-api-doc}), representing an
enormous amount of knowledge. In addition to simply reading these
articles, the articles contain a vast trove of information of interest
to researchers for machine aided questions (Kong \& Gerstein, 2018; Usai
et al., 2018). For example, many researchers are interested in
statistical outcomes of articles that can be extracted from numeric
results: P-values, effect sizes, means, and more. In addition,
researchers are often interested in words in articles, their use through
time, and the contexts they are found in.

Text-mining is the broad term associated with pulling information out of
articles. Given the importance of text-mining, good text-mining tools
are needed to make it easier for researchers to do. Graphical user
interface (GUI) based text mining tools are available (e.g., Ba \&
Bossy, 2016; Cañada et al., 2017; Muñoz, Kissling \& Loon, 2019) and
some research papers have used them (Chaix et al., 2019), but given the
urgent recent call to action for more reproducible research (Open
Science Collaboration, 2015; Camerer et al., 2016, 2018), we must move
away from GUI based tools as fast as possible. A number of examples of
programmatic tools can be found in the literature. For example, Sinclair
et al. (2016) present a tool in Python called seqenv for the domain
specific task of linking sequences to environments through text mining.

Most recent text-mining papers do not use programmatic approaches,
highlighting the need for more programmatic text-mining tools, and
increased discussion of those tools to increase awareness. For example,
many papers search Web of Science using their web interface, and
downloading papers manually (Ding, Li \& Fan, 2018; McCallen et al.,
2019). Many of these papers doing GUI based searching and paper
downloads are using R or Python downstream for analysis; replacing GUI
based data acquisition with programmatic approaches will improve
research.

The R programming language is used widely throughout many academic
fields; tools in R for text mining are of particular importance because
they can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those tools for
searching for, acquiring, and extracting parts of texts (e.g., title,
abstract, authors). Most of the packages presented here are part of the
rOpenSci suite (\url{https://ropensci.org/}).

@@

\begin{longtable}[]{@{}llll@{}}
\toprule
Data Provider & Content Type & rOpenSci Package &
Documentation\tabularnewline
\midrule
\endhead
Crossref & Metadata & rcrossref/crminer & \footnote{\url{https://api.crossref.org}}\tabularnewline
DataCite & Metadata & rdatacite & \footnote{\url{https://support.datacite.org/docs/api}}\tabularnewline
Biodiversity Heritage Library & Full content/Metadata & rbhl &
\footnote{\url{http://bit.ly/KYQ1Rd}}\tabularnewline
Public Library of Science (PLoS) & Full content/altmetrics & rplos &
\footnote{\url{http://api.plos.org/solr}}\tabularnewline
Scopus (Elsevier) & Full content/Metadata & fulltext & \footnote{\url{http://bit.ly/J9S616}}\tabularnewline
arXiv & Full content/Metadata & aRxiv & \footnote{\url{https://arxiv.org/help/api/index}}\tabularnewline
Biomed Central (via Springer) & Full content/Metadata & fulltext &
\footnote{\url{https://dev.springer.com/}}\tabularnewline
bioRxiv & Full content/Metadata & fulltext & \footnote{\url{http://www.biorxiv.org/}}\tabularnewline
PMC/Pubmed (via Entrez) & Full content/Metadata & rentrez & \footnote{\url{https://www.ncbi.nlm.nih.gov/books/NBK25500}}\tabularnewline
Europe PMC & Full content/Metadata & europepmc & \footnote{\url{https://azure.microsoft.com/en-us/services/cognitive-services}}\tabularnewline
Microsoft Academic Search & Metadata & microdemic & \footnote{\url{https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97}}\tabularnewline
Directory of Open Access Journals & Metadata & jaod & \footnote{\url{https://doaj.org/api/v1/docs}}\tabularnewline
JSTOR Data for Research & Full content & jstor & \footnote{\url{https://www.jstor.org/dfr/}}\tabularnewline
ORCID & Metadata & rorcid & \footnote{\url{https://pub.orcid.org/}}\tabularnewline
Wikimedia's Citoid & Citations & rcitoid & \footnote{\url{https://en.wikipedia.org/api/rest_v1/\#/Citation/getCitation}}\tabularnewline
Open Citation Corpus & Citations & citecorp & \footnote{\url{http://opencitations.net/}}\tabularnewline
Fatcat & Metadata & none & \footnote{\url{https://fatcat.wiki/}}\tabularnewline
SHERPA/RoMEO & Journal Level Metadata & rromeo & \footnote{\url{http://www.sherpa.ac.uk/romeo/apimanual.php?la=en\&fIDnum=\%7C\&mode=simple}}\tabularnewline
CORE & Full content/Metadata & rcoreoa & \footnote{\url{https://core.ac.uk/}}\tabularnewline
Dissemin & Metadata & dissemr & \footnote{\url{https://dissemin.readthedocs.io/en/latest/api.html}}\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{fatcat}{%
\subsection{Fatcat}\label{fatcat}}

Fatcat is a project from Ben Newbold of the Internet Archive Labs. It is
a ``versioned, publicly-editable catalog of research publications:
journal articles, conference proceedings, pre-prints, blog posts''.
Fatcat is currently does not have an R client, but is used inside of the
\href{https://github.com/ropensci/fulltext}{fulltext} package.

\hypertarget{sherparomeo}{%
\subsection{SHERPA/RoMEO}\label{sherparomeo}}

SHERPA/RoMEO (\url{http://sherpa.mimas.ac.uk/romeo/index.php})
aggregates and analyses publisher open access policies and provides
summaries of self-archiving permissions and conditions of rights given
to authors. The {[}rromeo{]}{[}{]} is an R interface to SHERPA/RoMEO.

\hypertarget{core}{%
\subsection{CORE}\label{core}}

CORE (\url{https://core.ac.uk/}) touts itself as the world's largest
collection of open access research articles, providing metadata on
journals and articles, as well as access to the full text of articles.
The \href{https://github.com/ropensci/rcoreoa}{rcoreoa} R package
interfaces with the CORE API.

\hypertarget{dissemin}{%
\subsection{Dissemin}\label{dissemin}}

Dissemin (\url{https://dissem.in/}) detects papers behind pay-walls and
invites authors to upload them to an open repository. Dissemin provides
metadata including links to open versions of articles. The
{[}dissemr{]}{[}{]} R package interfaces with the Dissemin API.

\hypertarget{fulltext-a-toolset-for-text-mining-in-r}{%
\section{fulltext: a toolset for text mining in
R}\label{fulltext-a-toolset-for-text-mining-in-r}}

\href{https://github.com/ropensci/fulltext}{fulltext} is a general
purpose R package for the data part of text-mining: search for articles,
get links to articles, get article abstracts, and fetch full text of
articles. The \texttt{fulltext} package is always adding additional data
sources as time allows (See Table 1). Starting from searching for
articles, the outputs of search can be fed into a function to get links
to those articles, or to get abstracts for those articles, or to fetch
their full text. The following is a breakdown of the major distinct
parts of \texttt{fulltext}.

@@

Text-mining will always be a complex task given all the layers involved:
often temporal time-span of research questions; varied permissions among
researchers and their articles they're trying to access; varied
approaches to getting full text (xml vs pdf vs plain text); and more.

Programmatic text-mining is a first step towards making text-mining
easier. The R ecosystem is an especially good place to do text-mining
because there are many packages for text-mining analysis, and endless
packages for any required statistical analyses. In addition, rOpenSci
and others are building up a set of packages in R for searching for and
acquiring full text programatically to help make the research workflow
as reproducible as possible.

Future work for \texttt{fulltext} includes:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Adding more publisher plugins
\item
  Fine tuned user control over publishers
\item
  Improve VPN/proxy controls
\item
  Incorporate more search engines to help resolve URLs for fulltext
  versions
\item
  Improve documentation
\end{enumerate}

With respect to what publishers can do to make text-mining easier,
publishers should:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  provide XML if they have it
\item
  not change URL patterns so often, or at all
\item
  maintain consistent URL patterns among journals, years, etc.
\item
  keep their Crossref metadata up to date
\item
  open up their citation data
\end{enumerate}

@@

\leavevmode\hypertarget{ref-Canada2017}{}%
Cañada A., Capella-Gutierrez S., Rabal O., Oyarzabal J., Valencia A.,
Krallinger M. 2017. LimTox: A web tool for applied text mining of
adverse event and toxicity associations of compounds, drugs and genes.
\emph{Nucleic Acids Research} 45:W484--W489.

@@

\leavevmode\hypertarget{ref-Ding2018}{}%
Ding Z., Li Z., Fan C. 2018. Building energy savings: Analysis of
research trends based on text mining. \emph{Automation in Construction}
96:398--410.

@@

\leavevmode\hypertarget{ref-McCallen2019}{}%
McCallen E., Knott J., Nunez-Mir G., Taylor B., Jo I., Fei S. 2019.
Trends in ecology: Shifts in ecological research themes over the past
four decades. \emph{Frontiers in Ecology and the Environment}
17:109--116.

\leavevmode\hypertarget{ref-Munoz2019}{}%
Muñoz G., Kissling WD., Loon EE van. 2019. Biodiversity observations
miner: A web application to unlock primary biodiversity data from
published literature. \emph{Biodiversity Data Journal} 7.

@@

\leavevmode\hypertarget{ref-Sinclair2016}{}%
Sinclair L., Ijaz UZ., Jensen LJ., Coolen MJ., Gubry-Rangin C.,
Chroňáková A., Oulas A., Pavloudi C., Schnetzer J., Weimann A., Ijaz A.,
Eiler A., Quince C., Pafilis E. 2016. \texttt{Seqenv}: Linking sequences
to environments through text mining. \emph{PeerJ} 4:e2690.
